Relationships in the extremes and their influence on competition and coexistence

### 2021-07-07 paper/Paper.Rnw
chose delta and mudif vals
further progress in methods and associated sup mat

\begin{document}
%\SweaveOpts{concordance=TRUE}

%The following is where you load in the numeric results that will be embedded in the text
<<eval=T,echo=F,message=F,warning=F>>=
system(paste("biber", sub("\\.Rnw$", "", current_input())))

#place R code here for loading in necessary variables from Results
#ordinarily this would be a readRDS command, but I'm just defining a variable here for now
allregres<-matrix(rnorm(10),2,5)
@

@@

Left- or right-tail associated noise $(b_1,b_2)$, with $b_i$ normally distributed with mean $\mu_i$ and standard deviation 
$\sigma$, was generated as pictured in Fig. \ref{fig:noise}, with details in SI section \ref{SI-sect:noise}.
To achieve goal G2 from the Introduction, i.e., to explore using a simple model the relative magnitudes of the contributions 
to coexistence that come from ATA versus other mechanisms, we plotted the terms $\Delta_i^0$, $\Delta_i^E$, $\Delta_i^C$, 
$\Delta_i^{(E\shp C)}$, $\Delta_i^{[E \nat C]}$ and $\Delta_i^{[EC]}$ [see (\ref{eq:full_decomp})] for the lottery model
against the independent parameters of the model, $\mu_1$, $\mu_2$, $\sigma$ and $\delta$; and then we compared their
values on the plots. Some mathematical shortcuts, summarized here, greatly simplified this plotting. First, we showed that, for the 
lottery model, none of the components in (\ref{eq:full_decomp}) depends independently on $\mu_1$ and $\mu_2$;
instead they all depend only on the difference $\mu_1-\mu_2$ (SI section \ref{SI-sect:math_simplifications}). So we only plotted 
against $\mu_1 - \mu_2$. Second, we showed that the values of all components in (\ref{eq:full_decomp}) were the same 
for both our left- and right-tail associated noises (SI section \ref{SI-sect:math_simplifications}). So we only generated plots 
for left-tail associated noise. For each combination of $\delta=W, X, Y, Z$
%***DAN: Jasmin, please select 4 sensible values, including \delta=1, and put the ones you select here
and $\mu_1-\mu_2=A,B,C,D$,
%***DAN: Jasmin, please select 3-4 sensible values and put the ones you select here
we plotted all the components of (\ref{eq:full_decomp}) against $\sigma$. Relative importance of contributions was judged from the 
plots to assess whether and to what extent ATA effects on coexistence were meaningfully large for the lottery model, compared to 
the contributions of other mechanisms.
%***DAN: Jasmin, if you select 4 values for mu1-mu2 and 4 for delta, then Fig. 3 is going to have 16 panels arranged in a 
%4 by 4 grid, each of which will plot all the components against sigma. You'll need to use different colors and/or line
%types for the components and have a legend.
The simulation methods of \cite{Ellner_2019} were used to compute the components of (\ref{eq:full_decomp}).
Computationally efficient means of carrying out these simulations, for the lottery model, are elaborated in 
SI section \ref{SI-sect:efficient_computation}.
%***DAN: Dan, you still need to write that sup mat section

@@

\begin{figure}
\caption{Pedagogical figure somehow displaying the different outcomes of competition in ATA and symmetric cases for an example.
This will be cited from the Intro, so needs to be understandable to someone who does not know any of the ideas we introduce yet.
So in particular you cannot use IGR as your metric of coexistence, you need to use something more intuitive. You will pick
model parameter values and do two sims, one with ATA noise and one with symmtric noise, and show different outcomes. We should
discuss and possibly take several iterations to get this figure right since it's a pedagogically important figure - if the reader
is not convinced in the Intro that we are studying something that makes a difference, they are not going to slog through
the Theory section, which is inevitably going to be tough going for many readers.\label{fig:pedagog}}
\end{figure}

>>>>

\begin{document}
\SweaveOpts{concordance=TRUE}
%\SweaveOpts{concordance=TRUE}

%The following is where you load in the numeric results that will be embedded in the text
%<<eval=T,echo=F,message=F,warning=F>>=
%system(paste("biber", sub("\\.Rnw$", "", current_input())))

%#place R code here for loading in necessary variables from Results
%#ordinarily this would be a readRDS command, but I'm just defining a variable here for now
%allregres<-matrix(rnorm(10),2,5)
%@

@@

Left- or right-tail associated noise $(b_1,b_2)$, with $b_i$ normally distributed with mean $\mu_i$ and standard deviation 
$\sigma$, was generated as pictured in Fig. \ref{fig:noise}, with details in SI section \ref{SI-sect:noise}.
To achieve goal G2 from the Introduction, i.e., to explore using a simple model the relative magnitudes of the contributions 
to coexistence that come from ATA versus other mechanisms, we plotted the terms $\Delta_i^0$, $\Delta_i^E$, $\Delta_i^C$, 
$\Delta_i^{(E\shp C)}$, $\Delta_i^{[E \nat C]}$ and $\Delta_i^{[EC]}$ [see (\ref{eq:full_decomp})] for the lottery model
against the independent parameters of the model, $\mu_1$, $\mu_2$, $\sigma$ and $\delta$; and then we compared their
values on the plots. Some mathematical shortcuts, summarized here, greatly simplified this plotting. First, we showed that, for the 
lottery model, none of the components in (\ref{eq:full_decomp}) depends independently on $\mu_1$ and $\mu_2$;
instead they all depend only on the difference $\mu_1-\mu_2$ (SI section \ref{SI-sect:math_simplifications}). So we only plotted 
against $\mu_1 - \mu_2$. Second, we showed that the values of all components in (\ref{eq:full_decomp}) were the same 
for both our left- and right-tail associated noises (SI section \ref{SI-sect:math_simplifications}). So we only generated plots 
for left-tail associated noise. For each combination of $\delta=0.2, 0.5, 0.8, 1.0$
and $\mu_1-\mu_2=0.0,-0.3,-0.6,-0.8$,
we plotted all the components of (\ref{eq:full_decomp}) against $\sigma$. Relative importance of contributions was judged from the 
plots to assess whether and to what extent ATA effects on coexistence were meaningfully large for the lottery model, compared to 
the contributions of other mechanisms.
%***DAN: If we select 4 values for mu1-mu2 and 4 for delta, then Fig. 3 is going to have 16 panels arranged in a 
%4 by 4 grid, each of which will plot all the components against sigma. We'll need to use different colors and/or line
%types for the components and have a legend. Sort it all out as one big multi-panel pdf fig, after running all the sims.
The simulation methods of \cite{Ellner_2019} were used to compute the components of (\ref{eq:full_decomp}).
Computationally efficient means of carrying out these simulations, for the lottery model, are elaborated in 
SI section \ref{SI-sect:efficient_computation}.

%What's next
%1) Write code implementing the above and making a figure - that'll be Fig. 3
%2) Get Jasmin to work on Fig. 1 and Fig. 2 (Fig. 2 is largely done). Fig 1 should be facilitated by the work to make Fig. 3 in the previous item.
%3) Decide what other results we will put in the main text and write a paragraph on methods for the other results (for both the main text
%and sup mat results), put it above.
%4) Think about this: Theoretical fig exploring why ATA matters to r_1\1 for the lottery model and hopefully more gernerally - try to get
%at the intuition here. See if you can edge to more generality here, or not?

%Ideas while driving to pick up Max:
%1) I think in every case for the lottery model, making tail associations asymmetric reduced the potential for coexistence - I think that's what Jasmin told me, but if we present it we should make sure I'm not remembering wrong, of course. Is that generally the case, for other models? What are the features of a model, or of the growth rate of the invader, that make it so that more asymmetric tail associations means reduced coexistence? Does this have implications for climate change?
%2) In this paper with Jasmin you need to cite the two papers you just got from the professor at University of Massachusetts, Lowell. I mean the papers about changes in the special extent of heat waves: https://agupubs.onlinelibrary.wiley.com/doi/full/10.1029/2020GL087097 and https://iopscience.iop.org/article/10.1088/1748-9326/ab4b41
%3) Discussion content: We should (in future) study tail associations between environmental variables, and how these might be changing. However, more than that, the variables E in the lottery model are not actual environmental variables but rather are the effects of environmental variables on the species of the model. Probably real environmental variables are actually related to vital rates in a non-linear and probably often threshold-like way. Threshold relationships can produce tail association, as we have showed elsewhere. Changes in the means of true environmental variables may change where those distributions of environmental variables are compared to thresholds, which may change the nature of tail associations between the effects of those environmental variables.

@@

\begin{figure}
\caption{Pedagogical figure somehow displaying the different outcomes of competition in ATA and symmetric cases for an example.
This will be cited from the Intro, so needs to be understandable to someone who does not know any of the ideas we introduce yet.
So in particular you cannot use IGR as your metric of coexistence, you need to use something more intuitive. You will pick
model parameter values and do two sims, one with ATA noise and one with symmtric noise, and show different outcomes. We should
discuss and possibly take several iterations to get this figure right since it's a pedagogically important figure - if the reader
is not convinced in the Intro that we are studying something that makes a difference, they are not going to slog through
the Theory section, which is inevitably going to be tough going for many readers.\label{fig:pedagog}}
\end{figure}
%possible examples:
%population simulations - shows symmetric with faster fluctatuations to extremes of coexistence while asymmetric has longer periods of noticeable coexistence and dominace/exclusion; may be unclear
